---
title: "Task 1 Analysis"
output: html_document
---

```{r setup}
source('../get_packages.R') # get all necessary packages

library(tidyverse)
library(readr)
library(plotly)

data_location = '//vwfs05/3BC-SHARE$/OST-R Priority Projects/16. Contributing Factors to Trends in Roadway Fatalities/Data/CrashData'

# data_file = 'AcVeUPD.csv'
# 
# d <- read_csv(file.path(data_location, data_file))

# Load from RData file, created previously from the csv. Much faster to load

load(file.path(data_location, 'AcVeUPD.RData'))

```

# Control loss

`grouping` is the main variable to look at. Grouping has 12 levels of pre-crash scenario.

- Do for FARS and GES/CRSS separately

- Outcome variable: count of crashes

- Predictor variables:  
  + `h_age_svCAT`: Driver age of subject vehicle, binned into categories ** Does include 1700 0 to 5 year old drivers... 
  + `speedrel_sv`: Speeding related **
  + `vsurcond_sv`: Surface condition for subject vehicle ** Dry, Slick (not Dry or Other), and Other + Reported as Unknown
  
- Mediator variables (possibly include): 
  + `h_mdlyr_sv`: Vehicle Model Year - no
  + `YEAR`: Year 
  + `p_inj_count_{A/B/C/DIED/K/O/U}`: Count or weight of number of persons injured on each level of KABCO scale
  + `vspd_lim_sv`: Posted speed limit for subject vehicle - no
  + `vtype_sv`: Vehicle body type for subject vehicle - no
  
- Grouping variables (probably don't include at first):
  + `scenario`: pre-crash scenario defiened by Volpe. 
  + `RUR_URB`: Rural or urban
  
  
  
```{r prep_fars}

counts_yr <- d %>%
  filter(grouping == 'b. Control Loss') %>%
  mutate(SOURCE2 = ifelse(SOURCE == 'FARS', 'FARS', 'GES/CRSS')) %>%
  group_by(SOURCE2, YEAR, 
           h_mdlyr_pv,
           speedrel_pv,
           h_age_pv,
           scenario) %>%
  summarize(case_count = n(),
            crash_count = sum(WEIGHT))


counts_yr <- d %>%
  filter(grouping == 'b. Control Loss') %>%
  mutate(SOURCE2 = ifelse(SOURCE == 'FARS', 'FARS', 'GES/CRSS')) %>%
  group_by(SOURCE2, YEAR) %>%
  summarize(case_count = n(),
            crash_count = sum(WEIGHT))

ggplot(counts_yr, aes(x = YEAR, y = crash_count)) +
  geom_point() + geom_line() +
  facet_wrap(~SOURCE2, scales = 'free_y') +
  ggtitle('Crash counts in total across years, by source')

```

```{r analysis1}

m1 <- lm()

```