---
title: "Variable Exploratory Analysis"
output: html_document
---

Exploratory data analysis of the OST-R Project 16 database

- Connect to the database snapshot in the Volpe shared drive
- Re-create some of Mikio's initial figures
- Make a slimmed-down version / specific tables for Tableau
- Save those in the shared drive

```{r setup}
source('../get_packages.R') # get all necessary packages

library(tidyverse)
library(readr)
library(plotly)

data_location = '//vwfs05/3BC-SHARE$/OST-R Priority Projects/16. Contributing Factors to Trends in Roadway Fatalities/Data/CrashData'

# data_file = 'AcVeUPD.csv'
# 
# d <- read_csv(file.path(data_location, data_file))

# Load from RData file, created previously from the csv. Much faster to load

load(file.path(data_location, 'AcVeUPD.RData'))

```

There are `r format(nrow(d), big.mark = ',')` rows and `r ncol(d)` columns of data.

TODO: 

- Get final list of column names and meta-data from Mikio:
  + Done, see Data Dictionary FINAL - 2020-07-17.xlsx.
  + Use Priority == P and _sv subject vehicle variables for initial Tableau work
  
- Create a reduced file for Tableau, with only key variables included


Notes from Mikio on the sample database:

>
> - This is a sample of the database we are creating (10,000 records for FARS, GES, CRSS)
> - The final database is >680,000 rows
> - The data contains 2 vehicles and relevant driver (these are named "variable"_1 or "variable"_2)
> "_nc" suffix denotes a variable that we tried to categorize based on its raw variable
> - This is a 1:1 of the raw data, case count wise. Not all variables are raw variables from the data, some are created by Volpe
> - This is created for our purposes and may not be the cleanest in its form, there are differences between N/A, null, "Missing", Unknown, > Not Reported, Blank, etc…


> The things you need to start with:
> - SOURCE
> - WEIGHT
> - YEAR
>
> 
> These 3 variables are the basis of the analysis, to identify trends over times in the variables. From here, its really plug and play. 
> 
> Happy to sit down and talk about anything (variable names, best next steps, etc…)
> My Task 1 Analysis are looking at these variables to refine our analysis, maybe we start in the same spot on how we show the data?
> ALCHL_IM, ALCOHOL, dr_dis'x'_nc, drimp'x'_nc, drvis'x'_nc, h_age, h_sex, highway_rc, TYP_INT,vtype
>

## Initial data exploration

```{r explore1}
# Plot FARS and GES/CRSS cases in total by year
table(d$SOURCE)

# Counts by year by source. Create a new two-level source variable to combine GES (2011 to 2015) with CRSS (since 2016)
# sum over WEIGHT to get the crash counts. Row sums = case counts

counts_yr <- d %>%
  mutate(SOURCE2 = ifelse(SOURCE == 'FARS', 'FARS', 'GES/CRSS')) %>%
  group_by(SOURCE2, YEAR) %>%
  summarize(case_count = n(),
            crash_count = sum(WEIGHT))


ggplot(counts_yr, aes(x = YEAR, y = crash_count)) +
  geom_point() + geom_line() +
  facet_wrap(~SOURCE2, scales = 'free_y') +
  ggtitle('Crash counts in total across years, by source')

```

## Grouping plots 

Aim to recreate slides 5 - 8

Validate against files in Teams folder
Refer to the `Analysis` sheet

- Pre-crash scenario plots: `grouping` 
- For FARS vs 
- Body Type



```{r grouping_tables}

table(d$grouping, d$SOURCE) # number of rows. Need to sum WEIGHTs by these categories to get crash counts for GES/CRSS

table(d$grouping, d$vtype_sv) # sv = subject-vehicle 

table(d$vtype_sv, d$vtype_pv) # sv = subject-vehicle, pv = principal-other vehicle (the vehicle being struck) 

# Getting towards the Danish plots
# table(d$grouping, d$vtype_sv, d$vtype_pv) # sv = subject-vehicle, pv = principal-other vehicle (the vehicle being struck) 

table(d$grouping, d$SOURCE)
```

### Slide 7: Proportion of crashes groups, over years

For FARS

```{r grouping_plots}

fars_total_yr <- d %>%
  filter(SOURCE == 'FARS') %>%
  ungroup() %>%
  group_by(YEAR) %>%
  summarize(sum_crashes_tot = sum(WEIGHT))

fars_grouping_yr <- d %>%
  filter(SOURCE == 'FARS') %>%
  ungroup() %>%
  group_by(YEAR, grouping) %>%
  summarize(sum_crashes_grp = sum(WEIGHT))

fars_grouping <- full_join(fars_total_yr, fars_grouping_yr)

fars_grouping_prop <- fars_grouping %>%
  ungroup() %>%
  group_by(YEAR, grouping) %>%
  summarize(prop_crashes_group = round(100 * sum_crashes_grp / sum_crashes_tot, 2))

slide7 <- ggplot(fars_grouping_prop, aes(x = YEAR, y = prop_crashes_group,
                               fill = grouping)) + 
  geom_bar(position = 'stack', stat = 'identity') +
  ylab('Proportion of crashes by pre-crash scenario') +
  xlab('Calendar Year')

slide7

# and adding an interactive version

ggplotly(slide7)


```


```{r save, include=FALSE}
# save(list = 'd',
#     file = file.path(data_location, 'AcVeUPD.RData'))
```